### 2015-05-20 graphs/openvpn-ftp-bench.tex
results: graphs

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 96.26)
        (aes256cbc:sha256, 89.29)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 46.01)
        (aes256cbc:sha256, 68.19)
    };
    \label{ba411e}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 7.16)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{out-of-tunnel}%"oot" for "out of tunnel"
\legend{software, ba411e, out-of-tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {file transfer over SSH tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 10.89)
        (aes256cbc:sha256, 8.19)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 10.67)
        (aes256cbc:sha256, 10.39)
    };
    \label{ba411e-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (none:none, 11.39)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{oot-tp}%"tp" for "throughput"
\legend{}
\end{axis}

>>>>

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 76.60)
        (aes256cbc:sha256, 76.03)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 83.74)
        (aes256cbc:sha256, 80.89)
    };
    \label{ba411e}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 7.16)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{out-of-tunnel}%"oot" for "out of tunnel"

\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 42.60)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {FTP transfer inside OpenVPN tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 4.78)
        (aes256cbc:sha256, 3.87)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 3.35)
        (aes256cbc:sha256, 2.84)
    };
    \label{ba411e-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (none:none, 11.39)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{oot-tp}%"tp" for "throughput"

\addplot[style={black,fill=RawSienna,mark=none}]
    coordinates {
        (none:none, 5.18)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

### 2015-05-20 results.tex
diffie-hellman
diagrams, implem drivers & OS theory
results: graphs

\chapter{Results}

@@

\section{TLS connections}

@@

>>>>

\chapter{Results and Analysis}\label{chap:results}

@@

\section{TLS connections}
% Begin with the benchmark done by Bastien on the raw number of verif/s with openssl.

This benchmark is done only with OpenVPN.
Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPSec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
As soon as the public key operations can be plugged into the crypto API, this use case should however be tested immediatly.

@@

\subsection{openSSH}

\begin{figure}[ht]
\input{graphs/openssh-bench}
\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}
\label{fig:openssh-bench}
\end{figure}

\subsection{OpenVPN}

\begin{figure}[ht]
\input{graphs/openvpn-ftp-bench}
\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}
\label{fig:openvpn-ftp-bench}
\end{figure}

We can see that adding a MAC computation aside the encryption merely lowers the performance when using the hardware.
Even though OpenSSL uses here an hihgly ARM-optimzed assembly implementation of SHA-256, it shows that the bottleneck is on the hardware side.

Indeed,

\subsection{IPSec}

\begin{figure}[ht]
\input{graphs/ipsec-ftp-bench}
\caption{FTP file transfer over an IPSec tunnel. The background stripped bars are the CPU usage.}
\label{fig:ipsec-ftp-bench}
\end{figure}

Some results have to be put into perspective with the fact that the implementation of SHA-256 is entirely C-based.
A more recent one using assembly instructions optimized for the NEON SIMD instruction set of the ARMv7 core could be used and would most probably yield better results.
The CPU usage of the software implementation would drop -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage could stay at the same level, we could expect a better throughput.

\subsection{Comparison}

\begin{figure}[ht]
\input{graphs/ftp-bench-comparison}
\caption{Comparison of file transfer methods. The background stripped bars are the CPU usage.}
\label{fig:ftp-bench-comparison}
\end{figure}

### 2015-05-20 technical-background.tex
rfc
results: graphs

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The RFC 2631~\cite{rfc2631} decribes the protocol:

\begin{enumerate}
	\item Alice generates once $p$ and $g$ (using precomputed parameters):
	\begin{description}[nosep]
		\item[p] large prime number
		\item[g] a generator of $\mathds{Z}_p^*$
	\end{description}
	\item Alice picks a random integer $x_a$ and computes $g^{x_a} mod\ p = y_a$.
	\item Alice sends $p$, $g$ and $y_a$ to Bob, signing everything using her private certificate.
	\item Bob checks the signature and picks $x_b$.
	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symetric key for further communications.
	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b} mod\ p$
\end{enumerate}

If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.
We will see in chapter~\ref{chap:results} that while a 1024-bit prime is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit primes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.

>>>>

\subsubsection{Diffie-Hellman}
Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symetric key during the following exchanges.
It uses the same kind of operation as RSA, that is modular exponentiation.
The protocol can be one of two type (\cite{rfc2631}, ...):
\begin{itemize}
	\item Static: the actors use their authenticated certificate to compute the shared secret.
	\item Ephemeral: the actors create a new pair of public/private keys from which the secret key is derived.
\end{itemize}
A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
Imagine that somehow, an opponent lays his hand on the shared secret.
If that secret has already been used, he can decipher all data transfered during past connections.
However, if the secret is new for every new connection, the compromission of the shared secret des not jeopardize past communications.
This is perfect forward secrecy: using a new key to protect the past.

Hereunder is the generation of an ephemeral shared secret.
For a static secret, Alice and Bob will simply use their static certificate, sparing one modular exponentiation.
\begin{enumerate}
	\item Alice generates once $p$ and $g$ (using precomputed parameters):
	\begin{description}[nosep]
		\item[p] large prime number
		\item[g] a generator of $\mathds{Z}_p^*$
	\end{description}
	\item Alice picks a random integer $x_a$ and computes $g^{x_a} mod\ p = y_a$.
	\item Alice sends $p$, $g$ and $y_a$ to Bob, signing everything using her private certificate.
	\item Bob checks the signature and picks $x_b$.
	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symetric key for further communications.
	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b} mod\ p$
\end{enumerate}

If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
However, any sane configuration will have both actors signing their ephemeral public key.
If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.
We will see in chapter~\ref{chap:results} that while a 1024-bit prime is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit primes.
Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.
